---
title: "Random Coefficient Logit estimation"
date: 2020-08-14
keep_tex: yes
output: 
  html_notebook: 
    code_folding: hide
    fig_caption: yes
    highlight: tango
    theme: united
    toc: yes
tags:
  - cool posts
  - category1
  - category2
header-includes: 
  - \usepackage{amsmath}
  - \usepackage{mathrsfs}
---

# Introduction

I provide here a relication in R of two important papers in the demand estimation literature. I use **BLPestimatoR**, a package created by Daniel Brunner to retrieve the original datasets from Berry Levinshon and Pakes (1995) and Nevo (2001).

```{r include = T, warning=FALSE, message = F}
rm(list = ls())
library(BLPestimatoR)
df1 <- productData_cereal
df2 <- productData_cars
summary(df2)
```

# General Model  

## Demand side

We model the demand as a function of product characteristics, heterogeneous consumer preferences and unknown parameters.Each consumer is assumed to purchase one unit of the good that gives her the highest utility.

$T$ market indexed by $t = 1, \ldots , T$ \
In each market there are $I_t$ consumers indexed by $i = 1, \ldots, I_t$ \
$J+1$ products indexed by $j = 0, \ldots, J$ \
$j=0$ is the outside option which is the situation when consumers decide not to purchase any brand. \

The conditional indirect utility is defined as:
$$
  u_{ijt} =\left\{
      \begin{aligned}
        &x_{j} \beta_i + \alpha p_{jt} + \xi_{jt} + \epsilon_{ijt} & \quad \text{if} \quad & j \in \{1,\ldots,J\}\\
        &\epsilon_{i0t} & \quad \text{if} \quad & j = 0 \\
      \end{aligned}
    \right.
$$
$x_{j}$ is a $(1 \times K)$ vector of observed product characteristics \
$p_{jt}$ is the price of a product $j$ in market $t$ ; $\alpha$ is the price valuation \
$\xi_{jt}$ is the mean valuation of the unobserved (by the econometrician) product characteristics \
$\epsilon_{ijt}$ is the error term, usually iid RV with a type I EV distribution \
$\beta_i$ is a $(K\times 1)$ parameter vector of random coefficients \

The random coefficient for characteristic $k$ is given by:
$$
\beta_{i}^{k} = \beta^{k} + \theta^{k} v_{i}^{k}
$$

$\beta^{k}$ is a mean valuation for characteristic $k$ \
$v_{i}^{k}$ is a random variable \
$\theta^k$ is the standard deviation across consumers (if $\theta^k$ = 0 $\rightarrow$ Logit) \

Conditional indirect utility becomes:
$$
u_{ijt} = \underbrace{x_{jt} \beta - \alpha p_{jt} + \xi_{jt}}_{\delta_{jt}} \:+ \: \underbrace{\sum_{k} x_{jt}^{k} \, \theta^{k} \, v_{i}^{k}}_{\mu_{jt}(v_i)} \: + \: \epsilon_{ijt}
$$

Note that if the consumer heterogeneity entered the model only through the separable additive random shocks $\epsilon_{ijt}$ and these shocks were Type 1 Extreme-Value distributed, the model would boil down to a simple Logit.

## Supply side

Following the notation of Nevo (2001). We have $F$ firms, each produces a subset $\mathscr{F}$ of $j=1,\ldots,J$ different brands. The profit of the firm f is then: 
$$ \pi_f = \sum_{j \in \mathscr{F} } (p_j - mc_j) M s_j (p) -C_f $$
where $s_j(p)$ is the market share of the brand $j$ as a function of all prices. $M$ is the size of the market. $C_f$ is the fixed cost of production.
$p_j$ satisfy the FOC:
$$ s_j(p) + \sum_{r \in \mathscr{F}} (p_r - mc_r) \frac{\partial s_{r}(p)}{\partial p_j} = 0 $$
Let's denote $S_{jr} = -\frac{\partial S_r}{\partial p_j}, \quad j,r = 1, \ldots, J$ and:

$$
\Omega_{jr}^{*} = \left\{
    \begin{array}{ll}
        1 & \mbox{if } \quad \exists f: \{r,s\} \subset \mathscr{F}_f \\
        0 & \mbox{otherwise}
    \end{array}
\right.
$$
$\Omega$ is a $(J \times J)$ matrix such as $\Omega_{jr} = \Omega_{jr}^{*} \times S_{jr}$.
The first order conditions can be rewritten:

$$s(p) - \Omega (p-mc) = 0 $$
$\Omega^{*}$ is called the ownership matrix and $\mathscr{F}_f$ is the ownership structure.

# Estimation   

## Estimation challenges

**Endogeneity problem** Price and the econometric error term are correlated. That is the reason we need to use instruments. Nevo (2001) proposes to use "Haussman instruments" which is the price of a brand in different cities within a region. Standard BLP instruments are the characteristics of products in other markets. They are correlated with price and are valid IV because we assume the choice of characteristics are exogeneous or at least predetermined.

## Estimation routine

First, we compute the mean utility levels $\delta_{.t}$ that solve the following system of equations:
$$ s_{.t} (x, p_ty,\delta_{.t};  \theta, \beta) = S_{.t} $$
where the predicted market shares $s_{.t}$ equates the observed market shares $S_{.t}$. The inversion is done numerically. \

Second, we compute the error term $\xi_{jt} = \delta_{jt}-x_{jt} \beta - \alpha p_{jt}$. \

Third, we exploit a moment condition that is the product of instrumental variables and structural error term $\xi$, to form a GMM estimator. Let define $Z = [Z_1, \ldots, Z_n]$ the set of instruments such that: $E[Z' \xi_j (\theta)] = 0$. The GMM estimator is:
$$ \hat{\theta} = \arg \min_{\theta} \quad \xi (\theta)' Z A^{-1} Z' \xi (\theta) $$
where $A$ is a consistent estimator of $E[Z' \xi \xi'Z]$. The optimal weight matrix under the assumption of homoscedasticity is $Z'Z$. Nevo (2001) proposes a two-step procedure. First, A is set to $Z'Z$, then the new estimate of the parameters $\theta$ are used to reconstruct the weight matrix:

$$ A = \frac{1}{n} \sum_{i=1}^{n} \xi (\theta)'\xi(\theta)Z'Z $$
I use a Quasi-Newton method with a user supplier gradient following Nevo (2001). I also try a Nelder-Mead non-derivative simplex method as in BLP. \

A note on standard errors: standard errors are computed blablabla...

**Using the fixed-point theorem** Blabla

**Using MPEC** Blabla



# Data

# Code

## Estimation steps

1. Prepare the data: draws from the distribution of $v$, run a TSLS estimation (Logit) to get initial values. <span style="color:red">RCOptimalInstruments.m</span> \
2. For a given value of $\theta$ and $\beta$, compute the market shares implied by the market share equation (simulation). <span style="color:red">ShareCalculation.m</span> \
3. For a given value of $\theta$, calculate the $\delta$ that equates the market shares computed in Step 2 to the observed market shares (contraction mapping/inner loop). <span style="color:red">delta.m</span> \
4. For a given $\theta$, compute the error terms $\xi_{jt}$ and then the value of the objective function. <span style="color:red">gmm.m</span> \
5. Compute the gradient matrix. <span style="color:red">jacob.m</span> \
6. Search for the value of $\theta$ that minimizes the objective function (outer loop) with steps 2-5 nested for every $\theta$ trial.<span style="color:red">RCOptimalInstruments.m</span> \
7. Compute standard errors for all parameters. <span style="color:red">seblp.m</span>
```{r include = T, warning=FALSE, message = F}
rm(list = ls())


```

